Practical Performance Evaluation of Entity Resolution Algorithms: Lessons Learned at PatentsView.org

### 2022-07-21 paper/main.tex
Update paper

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions to inventors, where a single person's name may be spelled in different ways on two applications, and where two different inventors with the same name may be difficult to distinguish. Much research has been done within the fields of economics, computer science and statistics to disambiguate inventor mentions \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

@@

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we have carried out a simple experiment which is described in detail in appendix \ref{appendix:example}. In short, we considered evaluating a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark data sets are obtained and used. In this experiment, we know that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

@@

@@

\begin{table}
    \label{tab:results}
    \centering
    \caption{\centering Estimated pairwise precision and recall (with estimated standard deviation) from our four benchmark datasets.}
    \begin{tabular}{lccc}
    \toprule
    \textbf{dataset} & \textbf{est. precision ($\hat \sigma$)} & \textbf{est. recall ($\hat \sigma$)} & \textbf{scope} \\
    \midrule
    Staff 1 & 86\% (5.6\%) & 94\% (2.1\%)  & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Staff 2 & 88\% (5.6\%) & 95\% (2.0\%) & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Israeli Benchmark & 79\% (NA) & 94\% (NA) & 1976 -- 1999 (U.S. granted)\\
    \cite{Li2014}'s Benchmark & 91\% (2.7\%) & 91\% (5.0\%) & 1976 -- 2010 (U.S. granted)\\
    \bottomrule
    \end{tabular}
\end{table}

@@

In our experiment, we have repeated 5,000 times the following 3-steps process:
\begin{enumerate}
    \item First, 100 (resp. 200 and 400) records were sampled and the ground truth clusters associated to them were recovered. This provided a ``benchmark" dataset which was used for evaluation.
    \item Second, a trivial precision estimate was obtained by computing precision over the benchmark dataset. That is, predicted cluster assignments were restricted to records which appear in the benchmark data and precision was compared for this. Most often than not, the result was an observation of $100\%$ precision.
    \item Third, we computed our proposed precision estimator which corresponds to lemma \ref{lemma:record_sampling_representation} and the estimator $\widehat P_{\text{block}}$ defined in equation \eqref{eq:P_block}, with blocks corresponding to clusters and with sampling probabilities $p_b \propto \lvert b \rvert$.
\end{enumerate}
The distributions of the these two precision estimates over the 5,000 repetitions and for sample of sizes 100, 200, and 400 are shown in figure \ref{fig:precision_problem}. Our proposed estimator is accurate and nearly unbiased, whereas the trivial precision estimates have almost nothing to do with actual algorithmic performance.

>>>>

For instance, the U.S. Patents and Trademarks Office (USPTO) makes available patent data dating back to 1790 \ob{(numerized full-text data is available from 1976)}. However, there is no standard for uniquely identifying inventors on patent applications. The result is a set of ambiguous mentions to inventors, where a single person's name may be spelled in different ways on two applications, and where two different inventors with the same name may be difficult to distinguish. \ob{Inventor mobility further complicates the use of contextual information for disambiguation.} Much research has been done within the fields of economics, computer science and statistics to disambiguate inventor mentions \ob{in this situation} \citep{trajtenberg2008identification, ferreira2012brief, ventura2013methods, Li2014, Ventura2015, kim2016random, yang2017mixture, morrison2017disambiguation, muller2017semantic, traylor2017learning, balsmeier2018machine,tam2019optimal, monath2019scalable, doherr2021disambiguation}.

@@

\begin{example}[Bias of precision computed on benchmark datasets]\label{first_example}
    To exemplify the problem with the trivial use of performance evaluation metrics on benchmark datasets, we have carried out a simple experiment which is described in detail in appendix \ref{appendix:example}. In short, we considered evaluating a disambiguation algorithm by sampling ground truth clusters and computing pairwise precision\footnote{Pairwise precision is the probability that a pair of record predicted to be a match is indeed a match. See section \ref{sec:proposed_estimators} for a formal definition} on this set of sampled clusters. This is analogous to the way that many real-world benchmark data sets are obtained and typically used. In this experiment, we know that the disambiguation algorithm has a precision of $52\%$ for the entire dataset.

@@

\begin{figure}[h]
    \centering
    \includegraphics{paper/clustering-example.pdf}
    \caption{\ob{Example of ground truth clustering $\mathcal{C}$ (represented by boxes with a full border) and predicted clustering $\widehat{\mathcal{C}}$ (rounded boxes with a dotted border) of elements $\mathcal{D} = \{1,2,\dots, 8\}$. Here $\mathcal{T} = \{(1,2), (2,3), (1,3), (4,5), (6,7)\}$ and $\mathcal{P} = \{(1,4), (2,3), (6,7), (7,8), (6,8)\}$. As such, $P = R = 2/5$ in this example.}}
    \label{fig:my_label}
\end{figure}

@@

\begin{table}
    \centering
    \caption{\centering Estimated pairwise precision and recall (with estimated standard deviation) from our four benchmark datasets.}
    \label{tab:results}
    \begin{tabular}{lccc}
    \toprule
    \textbf{dataset} & \textbf{est. precision ($\hat \sigma$)} & \textbf{est. recall ($\hat \sigma$)} & \textbf{scope} \\
    \midrule
    Staff 1 & 86\% (5.6\%) & 94\% (2.1\%)  & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Staff 2 & 88\% (5.6\%) & 95\% (2.0\%) & 1976 -- Dec. 31, 2022 (U.S. granted)\\
    Israeli Benchmark & 79\% (NA) & 94\% (NA) & 1976 -- 1999 (U.S. granted)\\
    \cite{Li2014}'s Benchmark & 91\% (2.7\%) & 91\% (5.0\%) & 1976 -- 2010 (U.S. granted)\\
    \bottomrule
    \end{tabular}
\end{table}

@@

In our experiment, we have repeated 5,000 times the following 3-steps process:
\begin{enumerate}
    \item First, 200 records were sampled and the ground truth clusters associated to them were recovered. This provided a ``benchmark" dataset which was used for evaluation.
    \item Second, a trivial precision estimate was obtained by computing precision over the benchmark dataset. That is, predicted cluster assignments were restricted to records which appear in the benchmark data and precision was compared for this. Most often than not, the result was an observation of $100\%$ precision.
    \item Third, we computed our proposed precision estimator which corresponds to lemma \ref{lemma:record_sampling_representation} and the estimator $\widehat P_{\text{block}}$ defined in equation \eqref{eq:P_block}, with blocks corresponding to clusters and with sampling probabilities $p_b \propto \lvert b \rvert$.
\end{enumerate}
The distributions of the these two precision estimates over the 5,000 repetitions are shown in figure \ref{fig:precision_problem}. Our proposed estimator is accurate and nearly unbiased, whereas the trivial precision estimates have almost nothing to do with actual algorithmic performance.
